Visualization and analysis of probability distributions of large temporal data

### 2022-02-24 report/response.Rmd

Not changed.
Figure is already published.

> Q2. Last sentence of Section 2.3.4: fix the table reference

Done.

> Q3. On page 23 the references to parts of Figure 2.4 need fixing.

Reference is in an acceptable format. No change as already published.

> Q4. In Figure 2.4, the outlier glyphs are in my opinion too big and dark and attract too much attention.
The log scale for the y-axis should be mentioned in the caption.

@@

Done. 
The outliers are made smaller and transparent and caption changed.

> Q5. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging. The plot deserves a bit more space.

Not changed as figure is already published. Furthermore, the figure already occupies one page.

@@

> Q1. On page 33 there is a mention of the hakear package, without a reference.

No change.
It appears in the abstract of the paper. For the thesis, I have added it since the abstract acts as a summary to the chapter. The package reference is mentioned in the Supplementary section of the paper. Moreover, the plan is to bring all the functions from `gravitas` to `hakear` when we publish it.

> Q2. In the Introduction state the methods are for continuous variables.

Done.
Added "for a continuous univariate dependent variable" in the last paragraph of Introduction of Chapter 3.

@@

> Q3. In Figure 3.1, the points are a bit hard to see, so it looks like whiskers extend beyond the data.

Done.
Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.2) and geom_jitter(alpha = 0.04 -> 0.03). Also jitter-width decreased to 0.2.

> Q4. On page 35, refer to months consistently using short or long names.

Done. x-axis labels changed to month-of-year for Figure 3.1 and 3.2.

@@

> Q8. Maybe include a reference for gestalt theory.

Done.

> Q9. In Section 3.2.2, I find the references to the null distribution confusing. Surely that is the design in
Figure 3.3(a) only?

Clarification: Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities. Data in Fig 3.3 (a) is consistent with the null distribution.

@@

> Q10. In table 3.1 $N_c$ is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
3.3.3) it is m.

Clarification: $N_c$ refers to the total number of contextual cyclic granularities. $m$ refers to the number of cyclic granularities we are considering together in the display. For example, contextual cyclic granularities could be $hour_{day}$, $day_{week}$ and $month_{year}$ and we want to visualize any one granularity at a time. So $N_c = 3$ and $m=1$.

Added $m$ to table 3.1  
$m$: number of cyclic granularities to display together

@@

> Q13. I like Figure 3.4. In (3), the dotted arcs only connect to a1 which might be misleading.

Caption changed to incorporate that.
Within-facet distances are illustrated in Panels 3) (when categories are un-ordered, shown only with respect to a1) and Panel 4) (when categories are ordered)

> Q14. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of observations across the levels of A and B? For example in Figure 3.3(d) if there are few values at xlevel=1 the comparison shown is less interesting

Clarification: As long as that there is enough data for each level or combination of levels, no adjustment for varying number of observations is done. If there are few data for one or more levels, subsequent categories could be collapsed to obtain precise estimates of quantiles.

@@

Added under the subsection "Characterizing distributions".

> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of $wpd_{perm}$ and on the residual definition at the bottom of the page. Use $\times$ for the equations.

@@

Done.

> Q16. Is there a practical reason why $wpd_{glm}$ is not working as expected for lower nx and nf?

Clarification: There is no practical explanation for this, but a more empirical one - with smaller $n_x$ and $n_f$, the variance in the values of $wpd_{raw}$ is greater. As a result, our modeling technique that attempts to fit $wpd_{raw}$ to the number of comparisons ($n_x\times n_f$) performs poorly. The increased volatility in the values of $wpd_{raw}$ with a smaller number of levels is due to the fact that no asymptotic distribution for the statistic "maximum" (which is the foundation of $wpd_{raw}$) across such small number of values could be produced.

> Q17. In the 3.3.1 algorithm, there is $m$, and then $M$.

@@

> Q18. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.

$m=1$ removed from the main chapter and included in supplements.

> Q19. The presentation of the material on the simulation study on page 48 could be improved. Where is the notation $wpd_{l,s}$ used? Figure 3.5 shows the m = 2 results only. The text alludes to a simulation involving different underlying distributions, but this is not mentioned again.

The simulation design and results corresponding to $m=2$ are presented. Similar design and results for $m=1$, although important, are not included in the main chapter but in the supplements. Simulation results involving different underlying distributions are also presented in the supplementary paper.

@@

> Q20. In Figure 3.5 the axis tick labels should be smaller. The blue and orange marks are hard to see. Maybe
show fewer panels?

Changed. Axis tick labels made smaller and space between facets minimised to allow for more space.

@@

> Q21. In Figure 3.6 the axis tick labels should be smaller. The blue and orange are hard to see. The caption
should refer to the rug. Maybe show fewer panels?

Changed. Axis tick labels made smaller and the rug plots are removed to make the display less busy.

@@

> Q22. In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend. Maybe use a different colour in the heatmap for the significant comparisons. State the threshold for significance in the caption. In Table 3.3 the caption should explain the threshold. (Maybe use colour instead of stars to indicate significance?)

CHANGED.
- Different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border    
- threshold explained in caption in Table 3.3   
- grey color explained in the text and caption

Clarification: Using color to convey significance instead of stars (in the Fig 3.3a) is a not a great idea because the varied shades of the tiles represent different values of the $wpd$. For example, both tiles may be significant, and their colours may differ to represent distinct $wpd$ values. To be consistent, Table 3.3 also has stars to represent significance.

@@

> Q23. The link in the Acknowledgements https://github.com/Sayani07/paper-hakear is not available. Neither are the supplementary materials. I understand this is for the paper version, I mention it for completeness.

The Github repository has now been made public. The link works now.

@@

> Q4. End of Chapter 4.1: incorrect reference to Section 2.7.

Done.

> Q5. I found Figure 4.1 very useful. In each box, would be helpful to put in “or”, in places where only one of the steps listed is performed, eq Normal quantile transform or Robust scaling. As there are many steps in the algorithm, it would be helpful to the reader to label the pipeline steps and to refer back to them in the text. The text in the Data pre-processing box does not make sense to me.

Rephrased the text in the  Data pre-processing section to make it easier to understand.

Figure 4.1 labels changed to make it consistent with the text.

@@

> Q6. The section describing RS and NQT is confusing. It states RS is applied to each time series separately. Is NQT also applied to each observation separately? How does “it could be useful to standardize it for
the selected set of significant granularities prior to computing the distances” relate to the following bullet points?

Clarification: Each observation (time series) is scaled independently in both JS-NQT and JS-RS. For the WPD approach, each observation might be scaled or not scaled, resulting in usable groups in either case.

> Q7. Page 64 “D is the Jensen-Shannon distance”.

No change. I had mentioned that $D$ stands for Jensen-Shannon distances.

@@

> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the distribution of each $v_t$.

No change. Stating the distribution of each $v_t$ is difficult because the levels for each granularity and design varies, therefore I chose to divide it into two tables.

> Q10. Where is $\mu$ in Table 4.2?

Clarification: $\mu$ is the difference between means considered for consecutive categories. It will vary depending on the chosen design and hence not tabulated.

@@

> Q11. In Section 4.3.3 what are the values of R and T?

Clarification: Various range of values of $R$ and $T$ are considered as stated in Table 4.1.

@@

> Q12. Figure 4.5 caption: MDS summary plots of what?

CLARIFICATION: MDS summary plots of clusters obtained from employing JS-NQT on the validation data sets
Added in the caption.

@@

> Q13. Figure 4.7 is missing labels (a), (b) (c).

Done. Added a, b representing 12 customers (rows) in each of them.

> Q14. Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?

Caption and text changed.  
The caption and description of the figure in the text have been changed to indicate that Fig. 4.9 depicts a visual summary of why three clusters are insufficient for this data set, however it is recommended by sindex summary plots. In a cluster analysis, it is standard practice to test a few cluster sizes before and after the optimal $k$ to see if a different number of clusters is useful. In this example, increasing the number of clusters actually unfolds some behavior which was not apparent with $k=3$ as suggested by sindex.

> Q15. The reference to the stationarity assumption in the Discussion needs Clarification.

Changed. "we have assumed the time series to be stationary" has been changed to "the conditional distributions are assumed to remain constant for the observation period" in the Discussion section.

@@

The ticmark labels are placed in both the figures but the houses are not brought to a same scale. More than comparing the two houses, it is important to compare the patterns of their distributions across months. They have different scales and showing them in the same scale is making the individual distributional differences seem less clear.

> Q4. In general, I like to see ticmark labels close to the extrema of the shown data points. So, in Figure 3.3a, there should be additional ticmark labels at -3 and +3, in 3.3b at -5
and +10, and so on.

Done.

> Q5. Before introducing a new distance measure in Section 3.2, I would like to see a literature review of existing distance measures and their limitations. Why is it necessary to introduce a new distance measure here? This is partially addressed on p. 41, but that should be placed earlier in the text. One clarifying question: Which density is represented by f in the equation? And shouldn't there be a dx at the end of the integral? Also, the distance measures that are mentioned at the end of that paragraph should be explained in one or two sentences.

Clarification 1:
The goal is to determine if there are any statistically significant differences between independent (unrelated) groups similar to a one-way or two-way ANOVA. A one-way ANOVA, for example, is analogous to $m=1$ (looking at one cyclic granularity at a time) in our approach. Assume we want to examine whether there is a significant difference in electricity demand on various days of the week. In this case, each day of the week may be regarded as an independent group, and we are looking at distributional differences rather than mean differences (as in a standard ANOVA). Similarly, considering $m=2$ (two cyclic granularities) is equivalent to a two-way ANOVA and is used to assess how distributions of quantitative variables vary according to the levels of two categorical variables rather than merely the mean or one measure of central tendency. There are other ways to obtain distributional differences, but we chose Jensen-Shannon distances between quantiles for ease of computation.

@@

> Q9. It would be helpful to add a specific link to a subsection in a cross-reference such as (See the supplements for more details.) on p. 42. Same on p. 47 and p. 48.

Mentioned which table or subsection needs to be referred and the link of the supplementary material provided.

> Q10. On p. 45 and in (3.2), it is not immediately clear whether nx; nf simultaneously have to be less than or equal to 5 so that $wpd_{perm}$ is being used - or whether only one of them has to be less than 5. Just a verbal clarification is needed. A practical scenario where this applies might be months (12) and weekdays/weekend (2).

Clarification: We have assumed $nx$; $nf$ simultaneously to be less than or equal to 5

@@

Done.

> Q13. The Results section and Figure 3.5 need some clarifications. The text states: Figure 3.5 shows that both the location and scale of the distributions change across panels. I suppose this figure relates to m = 2, but this is not stated in the main text. Table 3.2 lists both, m = 1 and m = 2. Moreover, in Figure 3.5, the y-axis ticmark labels overplot and the nx values are partially cut off.

Done.
We decided to remove the y-axis labels in the interest of space. The y-axis is showing the density values but we are only interested in its shape.

Clarification: Added $m=2$ in the caption of Figures 3.5 and 3.6.

@@

> Q17. In Figure 3.7a, it is not clear what the eight heatmaps represent. Do they belong to id 1, ..., id 8 ? Some explanation in the text or listing the id information on the left would be helpful. Similarly, on p. 53, the text states: id 7 and 8 have the same significant harmonies. I do not know where to look at the heatmaps in Figure 3.7a to find this information (and what is mentioned under items 2. and 3. in the text). Some explanation and labeling is needed here.

Caption changed to illustrate the plot better. Added box borders in (a) such that (a) and (b) display the household ids in the same order as indicated by the same color of the line plot in (b) and box border in (a).

> Q18. How were the harmony pairs selected and sorted in Figure 3.7 and Table 3.3? headmaps show 25 pairs, the parallel coordinate plot shows only 14 (same as Table 3.3). However, the last two are sorted differently, "hod wdwnd" (the first row in the table) is neither the first nor the last row in the parallel coordinate plot.

Clarification 1: Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities. The colors (in shades of red) represent the value of $wpd$ for each of the harmony pairs (in Table Table 3.3) and the grey tiles correspond to clashes. A darker shade of red corresponds to higher values of $wpd$.

Clarification 2: The harmony pairs in the parallel coordinate plot are arranged from highest to lowest values of $wpd$ averaged over all households and hence do not correspond to the first or last rows of Table 3.3.

@@

Caption rephrased to make it more clear.

For id: 7, there is more distributional difference between consecutive hours of the day compared to id:1. Also all the days of the week are different from each other in id:7. id:1 has similar behavior across all days within both weekdays and weekends. For this reason, the wpd for id:7 has a higher value than id:1 as the similar behavior within weekdays and weekends balances out the difference between weekdays and weekends. The display with hod vs wdwnd would have shown higher values of wpd for id:1 (as can be verified from Table 3).

> Q22. I didn't ask earlier, but how were the eight households for Section 3.4 being selected? Was this based on a random sample, or manually do obtain rather different energy usage patterns? This should be answered earlier on, but this also should be addressed in the Discussion: 
What can be expected when working with the full data set of Chapter 2, i.e., about 13,000. You indicate:
A future direction of work is to be able to explore and compare many individuals/subjects together for similar patterns across significant granularities. What would be computationally (and visually) feasible as 13,000 is a few orders of magnitude
higher than the number of eight households used in this example? Clearly, no full answer is expected here, but rather some speculation of what could possibly be done in the future (or what is done in the next chapter).

Clarification 1: The data for eight households for Section 3.4 are from friends and colleagues. They were manually chosen from friends with different profession and lifestyle to reflect different energy behavior.

Suppose there are $n$ observations. Our current clustering methodology has a time complexity of $O(n^2)$ since it calculates distance between all pairs. The time taken to compute distances between a pair of observations is also based on the number of cyclic granularities considered. For example, $1$, $2$ and $3$ cyclic granularities for JS-NQT are ~ 4.5s, 5.8s and 8.2s respectively. The times will be similar for JS-RS. For WPD, time taken will be more as the statistic $wpd$ need to be computed for each observation and each cyclic granularity. A rough estimate of working with $wpd$ working with two granularities with $(l,m)$ levels: ~(30s - 300s for 6<=l,m<=50) and (~40min - 50min for (l,m<=5). But this is a one-time job to obtain the significant granularities and can be run for a few individuals to get a sense of the significant granularities for the context. for example, for the electricity data, the granularities that were significant for most individuals were "wkdn/wday", "hour-of-day", "day-of-month" and hence clustering was run only on these variables.

@@

> Q9. Figure 4.7 speaks of (a) and (b) in the caption, but those letters do not appear in the figure. Either add them or refer to left and right. Moreover, in previous chapters, you used orange and green for the quartiles and 10th/90th quantile. Why not using the same colors and quantiles here? If this becomes too confusing for a reader, then at least, indicate in the caption which quantiles are covered by the gray areas.

Done.

> Q10. Be consistent across chapters. For example, in Table 3.3, you use wdwnd. On p. 72 and in Figure 4.7, you use wnwd. Adjust across all chapters. Check whether other abbreviations need to be standardized as well.

Done. Used "wnwd" through out Chapters 3 AND 4.

> Q11. Similar to Figure 4.9, it would help the reader that cluster P-3 is visually represented via three somewhat similar colors for Q-3, Q-4, and Q-5. Changing from red to blue/purple initially hides this information. Also arrange the legend from P-1 to P-3 and Q-1 to Q-5 to make it more obvious that P-1 & Q-1 and P-2 & Q-2 are identical.

### 2022-02-24 Rmd/04-gracsr.Rmd
fine tuning responses and incorporate missing comments

```{r mds-plot-validation, fig.cap = "MDS summary plots to illustrate the cluster separation for the range of mean differences (rows) under the different scenarios (columns). It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios. Between scenarios, separation is least prominent corresponding to Scenario (S3) where only one granularity is responsible for distinguishing the clusters."}

>>>>

```{r mds-plot-validation, fig.cap = "MDS summary plots of clusters obtained from employing JS-NQT on the validation data sets. It illustrates the cluster separation for the range of mean differences (rows) under the different scenarios (columns). It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios. Between scenarios, separation is least prominent corresponding to Scenario (S3) where only one granularity is responsible for distinguishing the clusters."}
